### 2017-03-10 manuscript/manuscript.tex
insignif
started intro

>>>>

The ecological condition of a stream or river, the life it supports, and the goods and services it provides, are influenced by the timing and magnitude of seasonal changes in water temperature. Temperature is a chief consideration in the management of fisheries, as it affects species distribution (Boisneau et al. 2008), life history (Elliott 1994), and migration timing (Boscarino et al. 2007). In the Puget Sound watershed of the American Pacific Northwest, [many species depend on temperature, and recreational and commercial fisheries depend on rivers maintaining temperautre within certain ranges]. Physically, temperature determines vapor pressure, surface tension, density (Stevens et al. 1975), gas solubility, reaction rates (Brezonik 1972) and more.  These properties also affect the efficacy of electrical power generation (Foerster & Lillestam 2010), and drinking water production (Ramaker et al. 2005).

These in turn are governed by climate at regional scales, and by physical attributes of the landscape at watershed or sub-watershed scales.

River networks, being fractal in structure, are naturally governed by environmental processes at multiple scales.  Water temperature is controlled at regional scales by climatic drivers, including air temperature and precipitation. At watershed or sub-watershed scales, physical attributes of the landscape, including geomorphology and geology, further structure temperature regimes, either directly or through interaction with climate drivers.

water temperature by air temperature, sediment influx by precipitation (Garner et al. 2013)—and at local scales by geographic, geologic, and anthropogenic effects.  Taken together, this hierarchical system is difficult to manage, as the physical characteristics of any one river may be largely the product of climatic drivers, while those of another may result primarily from within-watershed conditions. We argue that these characteristics, in particular water temperature and transported load, should be managed with specific attention to the relative influence of regional- and local-scale conditions for any given stream. The first aim of this chapter is to identify streams in the Puget Sound region whose temperature (°C), suspended solids (mg/L), and turbidity (NTU) are primarily predicted by regional temporal trends in air temperature and precipitation, and those that depart from regional patterns.  The second aim is to identify watershed features (land-use, geology, mean elevation, etc., specifics below) that correlate with such departures.

We hypothesize that:
1.  Air temperature will correlate strongly with water temperature for most streams (Ward 1985; Garner et al. 2013).  Deviations from this relationship will be strongest in those streams with glacially influenced hydrology (Lisi et al. 2015).
2.  Patterns in suspended load and turbidity, being highly dependent on soil type, lithology, and human activity, will be largely determined by within-watershed variables. Of these, we expect proportional agricultural cover to be most predictive (Waters 1995).
